\pkg{RProtoBuf}: Efficient Cross-Language Data Serialization in \proglang{R}

### 2014-01-23 papers/jss/article.Rnw
large scale -> large-scale protocol buffers -> Protocol Buffers
And remove some old commented out sections now.



git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@833 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@

%\title{RProtoBuf: Efficient Cross-Language Data Serialization in R}
%\author{by Dirk Eddelbuettel, Murray Stokely and Jeroen Ooms}

@@

%TODO(de) 'protocol buffers' or 'Protocol Buffers' ?
% MS: Lets standardize on 'Protocol Buffers'?

@@

%JO: can we just move the section 7 to here? It's only one paragraph%
%MS: I replaced section 7 with 2 sentences above.

@@

% The old section 7 in entirety:
%The \texttt{RProtoBuf} package uses the user defined tables framework
%that is defined as part of the \texttt{RObjectTables} package available
%from the OmegaHat project \citep{RObjectTables}.
%
%The feature allows \texttt{RProtoBuf} to install the
%special environment \emph{RProtoBuf:DescriptorPool} in the \proglang{R} search path.
%The environment is special in that, instead of being associated with a
%static hash table, it is dynamically queried by \proglang{R} as part of R's usual
%variable lookup. In other words, it means that when the \proglang{R} interpreter
%looks for a binding to a symbol (foo) in its search path,
%it asks to our package if it knows the binding "foo", this is then
%implemented by the \texttt{RProtoBuf} package by calling an internal
%method of the \texttt{protobuf} C++ library.

@@

One of the authors has used this design pattern for several large
scale studies of distributed storage systems \citep{sciencecloud,janus}.

@@

One example of a system that supports Protocol Buffers to interact
with \proglang{R} is OpenCPU \citep{opencpu}. OpenCPU is a framework for embedded statistical 
computation and reproducible research based on \proglang{R} and \LaTeX. It exposes a 
HTTP(S) API to access and manipulate \proglang{R} objects and allows for performing 
remote \proglang{R} function calls. Clients do not need to understand 
or generate any \proglang{R} code: HTTP requests are automatically mapped to 
function calls, and arguments/return values can be posted/retrieved
using several data interchange formats, such as protocol buffers.  
OpenCPU uses the \texttt{serialize\_pb} and \texttt{unserialize\_pb} functions
from the \texttt{RProtoBuf} package to convert between \proglang{R} objects and protobuf
messages. Therefore, clients need the \texttt{rexp.proto} descriptor mentioned
earlier to parse and generate protobuf messages when interacting with OpenCPU.

@@

The example above shows how the \texttt{HTTP GET} method retrieves a 
resource from OpenCPU, for example an \proglang{R} object. The \texttt{HTTP POST} 
method on the other hand is used for calling functions and running scripts, 
which is the primary purpose of the framework. As before, the \texttt{/pb} 
postfix requests to retrieve the output as a protobuf message, in this
case the function return value. However, OpenCPU allows us to supply the
arguments of the function call in the form of protobuf messages as well.
This is a bit more work, because clients needs to both generate messages 
containing \proglang{R} objects to post to the server, as well as retrieve and parse
protobuf messages returned by the server. Using Protocol Buffers to post
function arguments is not required, and for simple (scalar) arguments 
the standard \texttt{application/x-www-form-urlencoded} format might be sufficient.
However, with protocol buffers the client can perform function calls with
more complex arguments such as \proglang{R} vectors or lists. The result is a complete
RPC system to do arbitrary \proglang{R} function calls from within 
any programming language.

@@

% OpenCPU also provides a lot of meta-functionality such as handling
% of sessions, exceptions, security, and more. OpenCPU also makes it possible to store
% output of a function call on the server, instead of directly retrieving it. Thereby 
% objects can be shared with other users or used as arguments in a subsequent
% function call.
%  But in its essence, the HTTP API provides a simple way to perform remote 
% \proglang{R} function calls over HTTPS. The same request can be performed in Python as demonstrated
% below. The code is a bit verbose because to show how the REXP message is created from 
% scratch. In practice would probably write a function or small module construct a Protocol
% Buffer message representing an \proglang{R} list from a Python dictionary object. 
%
% \begin{verbatim}
% import urllib2;
% from rexp_pb2 import *;
%
% #create the post payload, i.e. list(n=42, mean=100)
% payload = REXP(
%  rclass = 5,
%    rexpValue = [
%      REXP(rclass = 2, realValue = [42]), 
%      REXP(rclass = 2, realValue = [100])
%    ],
%    attrName = [
%      "names"
%    ],
%    attrValue = [
%      REXP(rclass = 0, stringValue = [STRING(strval="n"), STRING(strval="mean")])
%    ]
%);
%
%#HTTP POST
%req = urllib2.Request(
%  "https://public.opencpu.org/ocpu/library/stats/R/rnorm/pb", 
%  data = payload.SerializeToString(), 
%  headers = {
%    'Content-type': 'application/x-protobuf'
%  }
%)
%res = urllib2.urlopen(req);
%        
%#parse output pb
%msg = REXP();
%msg.ParseFromString(res.read());
%
%#the return value is a double vector in this case
%print(msg.realValue);
%\end{verbatim}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
% Spelling Standardization:
% Protocol Buffers, not protocol buffers

@@

@@

@@

@@

@@

One of the authors has used this design pattern for several
large-scale studies of distributed storage systems
\citep{sciencecloud,janus}.

@@

One example of a system that supports Protocol Buffers to interact
with \proglang{R} is OpenCPU \citep{opencpu}. OpenCPU is a framework for embedded statistical 
computation and reproducible research based on \proglang{R} and \LaTeX. It exposes a 
HTTP(S) API to access and manipulate \proglang{R} objects and allows for performing 
remote \proglang{R} function calls. Clients do not need to understand 
or generate any \proglang{R} code: HTTP requests are automatically mapped to 
function calls, and arguments/return values can be posted/retrieved
using several data interchange formats, such as Protocol Buffers.  
OpenCPU uses the \texttt{serialize\_pb} and \texttt{unserialize\_pb} functions
from the \texttt{RProtoBuf} package to convert between \proglang{R} objects and protobuf
messages. Therefore, clients need the \texttt{rexp.proto} descriptor mentioned
earlier to parse and generate protobuf messages when interacting with OpenCPU.

@@

The example above shows how the \texttt{HTTP GET} method retrieves a 
resource from OpenCPU, for example an \proglang{R} object. The \texttt{HTTP POST} 
method on the other hand is used for calling functions and running scripts, 
which is the primary purpose of the framework. As before, the \texttt{/pb} 
postfix requests to retrieve the output as a protobuf message, in this
case the function return value. However, OpenCPU allows us to supply the
arguments of the function call in the form of protobuf messages as well.
This is a bit more work, because clients needs to both generate messages 
containing \proglang{R} objects to post to the server, as well as retrieve and parse
protobuf messages returned by the server. Using Protocol Buffers to post
function arguments is not required, and for simple (scalar) arguments 
the standard \texttt{application/x-www-form-urlencoded} format might be sufficient.
However, with Protocol Buffers the client can perform function calls with
more complex arguments such as \proglang{R} vectors or lists. The result is a complete
RPC system to do arbitrary \proglang{R} function calls from within 
any programming language.

@@
